---
title: "Co-piloting: how have we grown (dated `r sqlite_file_last_modified %>% format('%d %B %Y')`)"
author: "Lisa Hopcroft"
date: "`r lubridate::today()`"
output:
    bookdown::html_document2:
        number_sections: false
        toc: true
---

```{r setup, include=FALSE, echo=FALSE, warnings=FALSE}
library(tidyverse)
library(readr)
library(knitr)
library(ggplot2)
library(DiagrammeR)
library(glue)
library(flextable)
library(officer)
library(bookdown)
library(tibble)

knitr::opts_chunk$set( echo=FALSE, warnings=FALSE, message=FALSE, ft.align="left" )

theme_set(theme_minimal())

```

<style>
    p {margin: 10px 0px 25px; font-size:15px;}
</style>



```{r data_prep}

base_table = readRDS( here("dat", "base_table.Rds" ) )
jobs = readRDS( here( "dat","annotated_jobs.Rds") )
releases = readRDS( here( "dat", "annotated_releases.Rds" ) )
external_users_on_projects_table = readRDS( here( "dat", "external_users_on_projects_table.Rds" ) )
copilots = readRDS( here( "dat", "copilot_table.Rds" ) )

load( here("dat","report_variables.Rdat" ) )
load( here("dat","sqlite_variables.Rdat" ) )

```


```{r }

external_projects = base_table %>% filter( ! organisation_name %in% internal_organisations )
num_organisations = external_projects %>% pull( organisation_name ) %>% unique() %>% length()
num_projects = external_projects %>% pull( project_name ) %>% unique() %>% length() 

external_projects_ids = external_projects %>% pull( project_id ) %>% unique()
external_projects_list = external_projects %>% pull( project_name ) %>% unique()
external_workspace_list = external_projects %>% pull( workspace_name ) %>% unique()


```


# Quantitative description

The SQLite file was last modified on `r sqlite_file_last_modified`.
Other relevant start/end dates are shown in the table below.

```{r}
dates_to_present = tribble(
    ~variable, ~from, ~to,
    "Applications (submitted)", min( application_submitted_range ), max(application_submitted_range),
    "Applications (approved)", min( application_approved_range ), max(application_approved_range),
    "Workspaces", min(workspace_creation_range), max(workspace_creation_range),
    "Jobs (start)", min(job_start_range), max(job_start_range),
    "Jobs (end)", min(job_stop_range), max(job_stop_range),
    "Releases", min(release_range), max(release_range)
)
kable( dates_to_present )
```

## Overview: numbers of organisations, projects, copilots etc

```{r}
copilot_list = copilots %>% pull( copilot_name ) %>% unique()
num_copilots = copilot_list %>% length
```

We have `r num_copilots` active co-pilots:

```{r results='asis'}
cat(paste('-', copilot_list %>% sort), sep = '\n')
```


```{r}

external_users = external_users_on_projects_table %>% left_join(
    users_and_organisations, by="user_name"
) %>% filter( ! type %in% internal_organisations   )

returning_external_users = external_users %>%
group_by( user_name ) %>%
summarise( n_projects = project_id %>% unique %>% length )  %>%
filter( n_projects > 1 ) %>%
arrange( -n_projects )

```

We have `r external_users %>% pull(user_name) %>% unique() %>% length`
external users. `r returning_external_users %>% nrow` of these users
have worked on more than one OpenSAFELY project.

```{r}
column_name_map = list(
    organisation_name = "Organisation",
    project_name = "Project",
    project_count = "Number of projects",
    workspace_count = "Number of workspaces",
    total_num_files_released = "Number of files released",
    #overall_time_to_release = "Time to first release (in days)",
    #release_within_hard_limit = glue("Release within {hard_release_limit} days"),
    #release_within_soft_limit = glue("Release within {soft_release_limit} days"),
    #release_within_longer_limit = glue("Release within {longer_release_limit} days"),
    #release_within_no_limit = glue("Release (no limit)"),
    #n_hard_release = glue("Number released within {hard_release_limit} days"),
    #perc_hard_release = glue("Percentage released within {hard_release_limit} days"),
    #n_soft_release = glue("Number released within {soft_release_limit} days"),
    #perc_soft_release = glue("Percentage released within {soft_release_limit} days"),
    #n_longer_release = glue("Number released within {longer_release_limit} days"),
    #perc_longer_release = glue("Percentage released within {longer_release_limit} days"),
    n_projects_total = glue( "Total number of projects" ),
    copilot_name = "Co-pilot name",
    days_since_project_created = "Number of days since project was created",
    days_since_first_workspace = "Number of days since first workspace was created",
    project_creation_date = "Date that the project was created",
    first_workspace_start_date = "Date on which the first workspace was created",
    application_submitted_at = "Date on which the application was submitted",
    application_approved_at = "Date on which the application was approved",
    project_status = "Status of project (set by pilot)"
)

```

```{r}

projects_assigned_copilots = external_projects %>% filter( copilot_name != "None provided" )

num_projects_assigned_copilots = projects_assigned_copilots %>%
    pull( project_name ) %>% unique() %>% length()

external_releases = releases %>%
    filter( !is.na( release_created_at )) %>%
    filter( project_name %in% external_projects_list ) %>%
    ungroup( ) %>%
    mutate( specific_time_to_release = difftime( release_created_at, first_workspace_start_date ) )

projects_with_outputs = external_releases
    
num_projects_with_outputs = projects_with_outputs %>%
    pull( project_name ) %>% unique() %>% length()

projects_without_outputs = external_projects %>%
    anti_join( projects_with_outputs, by="project_name" )
num_projects_without_outputs = projects_without_outputs %>%
    pull( project_name ) %>% unique() %>% length()

```

Find some overall counts in Table \@ref(tab:overallcounts) below.

```{r}
overall_counts = tribble(
    ~count, ~n,
    "Total number of projects", num_projects,
    "Projects assigned co-pilots", num_projects_assigned_copilots,
    "Projects with outputs", num_projects_with_outputs, 
    "Projects without outputs", num_projects_without_outputs,
    "Number of active copilots", num_copilots, 
   " Number of organisations with pilots", num_organisations 
)

overall_counts %>%
    ### Format table
    flextable( ) %>%
    set_header_labels( values=list(count = "" ) ) %>%
    autofit() %>%
    theme_zebra() %>%
    footnote( i = 5, j = 1,
            value = as_paragraph("We have at least 1 former co-pilot that is not included in the list above"),
            ref_symbols = "a" ) %>%
    set_caption( glue( "Overall counts (as of {sqlite_file_last_modified})" ),
                autonum = run_autonum(seq_id = "tab", bkm = "overallcounts")) 

```


Projects may consist of multiple workspaces. A count of projects and workspaces per
organisation is provided in the Table \@ref(tab:numPKperorg).

```{r}

external_projects %>%
    ### Prepare data
    group_by( organisation_name ) %>%
    summarise( project_count = length(unique(project_name)),
                workspace_count = length( unique(workspace_name) ) ) %>%
    arrange( -project_count, organisation_name )%>%
    janitor::adorn_totals() %>%
    ### Format table
    flextable( ) %>%
    set_header_labels( values=column_name_map ) %>%
    autofit() %>%
    theme_zebra() %>%
    set_caption( "Number of projects and workspaces per organisation",
        autonum = run_autonum(seq_id = "tab", bkm = "numPKperorg")) 

```
